---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# MigStat

The goal of MigStat is to ease working with Migration Statistics
(Wanderungsstatistik). It's main functionalities are:

- prepare data to enable the fitting of spatial origin-destination
  flow models using
  [spflow](https://cran.r-project.org/web/packages/spflow/index.html)
- easy join of regional characteristics from the
  [inkar](https://www.inkar.de/) data
- generate random (currently only uniform) moves.
- make visualizations easy by:
  - creating maps with arrows to indicate destinations and size of
    migration flows from origin
  - calculate migration wins and losses for regions
  - calculate bivariate flows ("Umz端ge") to create [Circular Migration
    Flow
    Plots](https://www.r-bloggers.com/2014/03/circular-migration-flow-plots-in-r/)


## Installation

You can install the development version of MigStat from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("KonstantinBASachsen/MigStat")
```
# General Overview

The package evolves around three data sets:
- [Migration
  Statistics](https://www.forschungsdatenzentrum.de/de/10-21242-12711-2013-00-00-1-1-0)
  The example data set gives an idea about the data structure and what
  observations can be expected. Unfortunately there are only 200
  moves. For testing of visualizations and models this is not enough
  in my opinion. ^[The data set is actually not really needed. What is
  needed are the column names and some missings here and there. I am
  not sure if it is allowed to share the data structure publicely. I
  will find out and if so, I remove the dependence on the example data
  and provide a function to recreate it.]
- [inkar](https://www.inkar.de/) This data has very many interesting
  regional characteristics for many years and on different
  administrative levels like federal states, districts and
  municipalities.
- [shapefile](https://daten.gdz.bkg.bund.de/produkte/vg/vg250-ew_ebenen_1231/2014/vg250-ew_12-31.utm32s.shape.ebenen.zip)
  (2014 is not correct, 2013 should be used) This is used to obtain
  the geometry information of the regions and the population sizes.

## Examples

Currently the examples assume the user has access to the example
Migration Statistics (Beispieldaten der Wanderungsstatistik). So you
can not actually run the examples. It is merely intended to show the
workings.


### Example load data and join administrative units


```{r package_and_data, eval = TRUE}

library(MigStat)

ex_dat <- MigStat:::read_examples() ## reads shapefile and example migration statistics
shps <- ex_dat$shps
inkar_csv <- "/home/konstantin/Downloads/inkar_2021.csv"
inkar <- MigStat::read_inkar(inkar_csv)

```

Next we want to generate random moves between regions. First we
specify for which regions. Moves can be drawn between origin and
destination. Both can be of one of the following:
- federal states (st)
- districts (di)
- municipalities (mu)

In this case we generate moves where origin and destination are both
federal states. Currently only uniformly distributed moves are
created. This means every origin-destination pair has the same
probability that somebody moves between them. This is very
unrealistic.

```{r random_moves}

us <- "st" ## "unit_simple" on of c("st", "di", "mu")
dt <- MigStat::n_new_rows(dt = ex_dat$mig, shps = shps, us_o = us, us_d = us, n = 1000) # code should be vectorized, takes unnecessary long to run

```
The data is organized such that every row represents one move:

```{R data_structure}

dt

```

We are interested in the number of people moving between regions. The
distance between the centroids of regions is reported as well.

```{R}

flows <- get_flows(dt, shps, us = "st")
flows
```

### Create circular migration flow plot

```{r circ_plot}

migest::mig_chord(flows[, 1:3]) ### for this names of regions would be nice

```

### Create Arrow Plot

Now we can create an arrow plot that shows where and how many people
move from origin. First we extract from the shapefile the data we need.

```{R arrows_new}

shp <- MigStat:::clean_shp(shps = shps, us = us) # grabs the correct administrative column and drops unnecessary columns

name <- "Baden-W端rttemberg" ## name of origin region
dtarrow <- MigStat::get_arrow_data(flows, shp, name)
dtarrow
MigStat::arrow_plot(shp, dtarrow) ## draw plot, moves from BaW端 to BaW端 are a bit hidden

```

## Spatial Origin-Destination Flow Model

### Join covariates

Next we want to model migration flows. First we want to add some
covariates from the inkar data. The inkar data has `length(indic)`
different regional characteristics

```{R join_inkar}

zb <- "2013" ### level of covariates for 2013
indic <- unique(inkar[, Indikator]) ### these are all available covariates
idx <- c(428, 86, 344, 196, 193, 419, 240, 161, 153, 480)
vars <- indic[idx] 
shp <- MigStat::join_inkar_vars(shp = shp, inkar = inkar, vars = vars, us = us, zb = zb)

```

Now we looked a bit at the data; the data is in the right format and
we joined regional characteristics to be able to easily include
covariates to our model. This is essentially what the package is made
for. 

The next steps rely on the
[spflow](https://cran.r-project.org/web/packages/spflow/index.html)
package to build and estimate a spatial origin-flow model. This code
is not part of MigStat. It is merely intended to show how one can
proceed with data analysis.

### Create spatial weight matrix

```{r weight_matrix}

set_geom <- function(dt, geom_only = T) {
    dtgeom <- sf::st_set_geometry(dt, dt[, geometry])
    if (geom_only == TRUE) {
        dtgeom <- sf::st_geometry(dtgeom)
    }

    return(dtgeom)
}

mid_points <- sf::st_point_on_surface(set_geom(shp))
shp_nb <- list(
    "by_contiguity" = spdep::poly2nb(set_geom(shp)) ##,
##    "by_distance" = spdep::dnearneigh(mid_points,d1 = 0, d2 = 1500), ## 
##    "by_knn" = spdep::knn2nb(spdep::knearneigh(mid_points, 3))
)

plot(set_geom(shp))
plot(shp_nb$by_contiguity, mid_points, add = TRUE)
title("Contiguity") 

```


For efficient estimation of the model, the following data structure is
needed:

```{R model_structure}

idx <- !(colnames(shp) %in% c("geometry", "GEN"))
shp <- shp[, ..idx]
idx <- vars %in% colnames(shp) ## because not all desired vars could be joined we check if they are indeed part of shp
m_cols <- c("EWZ", vars[idx]) ## columns included in model
shp[,  (m_cols) := lapply(.SD, scale), .SDcols = m_cols]

shp_net <- 
  spflow::sp_network_nodes(
    network_id = "states",
    node_neighborhood = spdep::nb2mat(shp_nb$by_contiguity),
    node_data = shp,
    node_key_column = "AGS")


shp_net_pairs <- spflow::sp_network_pair(
  orig_net_id = "states",
  dest_net_id = "states",
  pair_data = flows,
  orig_key_column = "origin",
  dest_key_column = "destination")

shp_multi_net <- spflow::sp_multi_network(shp_net, shp_net_pairs)

```
